### 2016-07-13 NLO.tex
K factors in thesis

\subsection{The Squark Production Cross Section at Next-to-Leading Order}
The squark production cross section at next-to-leading order is composed out of three contributions. These are the Born (tree-level) cross section, the virtual and the real corrections. 
\begin{align}
\sigma^{\mathrm{NLO}} = \sigma^{\mathrm{B}} + \sigma^{\mathrm{V}} + \sigma^{\mathrm{R}}\label{eq:BVR}
\end{align}
All three contributions have to be calculated with next-to-leading order parton density functions and $\alpha_s$. There are five mass scales: the squark mass $m_{\tilde{q}}$, the gluino mass $m_{\tilde{g}}$, the center of mass energy of the colliding protons $\sqrt{S}$, the top-quark mass which is fixed to $m_t = \unit[172]{GeV}$ and the mass of the pseudoscalar $m_{\sigma}$. The next section summaries how the calculation has been automated within a \texttt{C++} program.

@@

\subsubsection{Remark on Prefactors in the Calculation}

@@

\caption{Dependence of the $K$-factor for the process $uu \to \tilde{u}_L\tilde{u}_R$ in the MRSSM on the squark mass $m_{\tilde{q}}$ for fixed $m_{\tilde{g}} = \unit[2000]{GeV}$ (top) and on the gluino mass $m_{\tilde{g}}$ for fixed $m_{\tilde{q}} = \unit[1000]{GeV}$ (bottom). The other parameters are $\sqrt{S} = \unit[13]{TeV}$, $m_{\tilde{\sigma}} = \unit[5000]{GeV}$ and $\mu_R = \mu_F = m_{\tilde{q}}$. The used parton density function are the one from table \ref{tab:MRSSMKfactors}. The missing smoothness of the plot originates from the uncertainty of the integration.}\label{fig:1LXsection_fixed_m_MRSSM}

@@

\caption{$K$-factors of the process $pp \to \tilde{u}_L\tilde{u}_R$  in the MRSSM for a selected set of masses. The center-of-mass energy is $\sqrt{S} = \unit[13]{TeV}$ and the pseudoscalar mass is fixed to \mbox{$m_{\mathrm{\sigma}} = \unit[5000]{GeV}$}. For $\sigma^{\mathrm{LO}}$ parton density functions \texttt{MMHT2014}(LHAPDF ID: $25000$) has been used, whereas for $\sigma^{\mathrm{NLO}}$ \texttt{MMHT2014}(LHAPDF ID: $25100$) was used.}\label{tab:MRSSMKfactors}

@@

It can be seen that the $K$-factors depend strongly on the ratio $\frac{m_{\tilde{g}}}{m_{\tilde{q}}}$ but not that strongly on the masses itself.\\

@@

\subsubsection{$K$-factors for Squark Production in the MSSM}
To compare the results of $K$-factors in the MRSSM to those in the MSSM, the cross sections in question have been calculated using the program \texttt{MadGraph5\_aMC@NLO} with the same parton density functions as before, i.e. \texttt{MMHT2014}. When generating the code for the pertaining processes, e.g. two protons to $\tilde{u}_L + \tilde{u}_R$, with
\begin{lstlisting}[style=Mybash]
generate p p > ul ur [QCD] $$go
\end{lstlisting}
the $s$-channel gluinos were omitted, with the addendum \texttt{\$\$go} as \texttt{MadGraph5} does not deal with divergences coming from the gluino in the on-shell limit. 
\begin{figure}[!htpb]
\begin{center}
\includegraphics[scale=.47]{figures/MSSM_uu_susu_Kfactors_msg=2000GeV.png}
\includegraphics[scale=.47]{figures/MSSM_uu_susu_Kfactors_msq=1000GeV.png}
\caption{Dependence of the $K$-factor for the process $pp \to \tilde{u}\tilde{u}$, where $\tilde{u}\tilde{u} \in \left\{ \tilde{u}_L\tilde{u}_R, \tilde{u}_L\tilde{u}_L, \tilde{u}_R\tilde{u}_R \right\}$, in the MSSM on the squark mass $m_{\tilde{q}}$ for fixed $m_{\tilde{g}} = \unit[2000]{GeV}$ (top) and on the gluino mass $m_{\tilde{g}}$ for fixed $m_{\tilde{q}} = \unit[1000]{GeV}$ (bottom). The parameters are chosen like in fig. \ref{fig:1LXsection_fixed_m_MRSSM}.The $K$-factors have been broken down to contributions coming from $pp \to \tilde{u}_L\tilde{u}_R$ and from $pp \to \tilde{u}_L\tilde{u}_L$. The latter equals the contributions coming from $pp \to \tilde{u}_R\tilde{u}_R$. The missing smoothness of the plot originates from the uncertainty of the integration.}\label{fig:1LXsection_fixed_m_MSSM}
\end{center}
\end{figure}
As already discussed, these contributions are only of sub percent order when $m_{\tilde{q}} > m_{\tilde{g}}$\cite{Gavin:2013kga} but increase for $m_{\tilde{g}} > m_{\tilde{q}}$ as indicated in table \ref{tab:MSSMMadGraphProspino}, which compares $\sigma^{\mathrm{NLO}}$ obtained using \texttt{MadGraph5} and the program \texttt{Prospino2}\cite{Beenakker:1996ed} for both scenarios.

\begin{table}[H]
\begin{center}
\begin{tabular}{c|c?c?c}
$m_{\tilde{q}}$ in GeV & $m_{\tilde{g}}$ in GeV & $\sigma^{\mathrm{NLO}}_{\texttt{MadGraph}}$ in fb & $\sigma^{\mathrm{NLO}}_{\texttt{Prospino2}}$ in fb \\
\hlinewd{2pt}
$1500$ & $1000$ & $9.49 \pm 0.03$ & $9.48 \pm 0.09$ \\
$1000$ & $2000$ & $13.32 \pm 0.05$ & $13.89 \pm 0.11$ 
\end{tabular}
\caption{$\sigma^{\mathrm{NLO}}(pp \to \tilde{u}_L\tilde{u}_R)$ in the MSSM obtained from \texttt{MadGraph5} and \texttt{Prospino2} for a selected set of masses. The center-of-mass energy is $\sqrt{S} = \unit[13]{TeV}$. The stated uncertainties originate from the phase space integration. Within the \texttt{MadGraph5} calculation, the $s$-channel gluino diagrams have been discarded. This becomes relevant only when $m_{\tilde{g}} > m_{\tilde{q}}$.}\label{tab:MSSMMadGraphProspino}
\end{center}
\end{table}

@@

\begin{table}[H]
\begin{center}
\begin{tabular}{c|c?c|c|c?c|c|c}
\multicolumn{2}{c?}{} & \multicolumn{3}{c?}{$p + p \to \tilde{u}_L + \tilde{u}_R$} & \multicolumn{3}{c}{$p + p \to \tilde{u} + \tilde{u}$} \\
\hlinewd{2pt}
$m_{\tilde{q}}$ in GeV & $m_{\tilde{g}}$ in GeV & $\sigma^{\mathrm{LO}}$ in fb & $\sigma^{\mathrm{NLO}}$ in fb & $K$ & $\sigma^{\mathrm{LO}}$ in fb & $\sigma^{\mathrm{NLO}}$ in fb & $K$\\
\hlinewd{2pt}
$500$ & $500$ & $966.4$ & $1328$ & $1.37$ & $2333$ & $3102$ \ignore{1328+2*887} & $1.33$\\
$500$ & $1000$ & $303.4$ & $352$ & $1.16$ & $1271$ & $1393$ \ignore{353 + 2*520} & $1.10$\\
$1000$ & $1000$ & $42.63$ & $58.5$ & $1.37$ & $122.2$ & $161.3$ \ignore{58.5+2*51.4} & $1.32$\\
$1000$ & $2000$ & $11.56$ & $13.3$ & $1.15$ & $66.36$ & $72.1$ \ignore{13.3+2*29.4} & $1.09$\\
$1500$ & $3000$ & $0.9166$ & $1.05$ & $1.14$ & $7.045$ & $7.57$\ignore{1.05+2*3.26} & $1.07$
\end{tabular}
\caption{$K$-factors of the process $pp \to \tilde{u}_L\tilde{u}_R$ and $pp \to \tilde{u}\tilde{u}$, where $\tilde{u}\tilde{u} \in \left\{ \tilde{u}_L\tilde{u}_R, \tilde{u}_L\tilde{u}_L, \tilde{u}_R\tilde{u}_R \right\}$, in the MSSM for a selected set of masses. The center-of-mass energy is $\sqrt{S} = \unit[13]{TeV}$.}\label{tab:KfactorsMSSM}
\end{center}
\end{table}

Again, it is clearly visible that the $K$-factors depend rather strongly on the ratio $\frac{m_{\mathrm{g}}}{m_{\mathrm{q}}}$ but not so much on the masses itself. Comparing with the results from the MRSSM in table \ref{tab:MRSSMKfactors}, one can state that the $K$-factors of the MRSSM are slightly larger. This might be the case, as there are additional degrees of freedom in the $R$-symmetric model, namely the octino which enters the gluino self-energy contributions (see second diagram from the right in \ref{fig:GluinoSE}) and sgluons which enter box diagrams (see lower right diagram in \ref{fig:1loopdiagrams}). The dependence on the sgluon mass is discussed within the next section.\\
The uncertainties of next-to-leading order cross sections, due to the phase space integration, within this section are of percent order. The uncertainty which corresponds to the scale variation does not increase 15\%.

\subsection{The Cross Section in the Limit of Large Sgluon Masses}
The cross section for squark production does not exist in the limit of an infinitely large sgluon mass, instead it was found that it diverges logarithmically:
\begin{align}
\lim_{m_{\sigma}\to\infty} \sigma(qq \to \tilde{q}\tilde{q}) \sim \ln \frac{m_{\sigma}^2}{\mu^2}.
\end{align}

\begin{figure}[H]
\begin{center}
\includegraphics[scale=.5]{figures/MRSSM_uu_susu_Kfactors_msq=1000GeV_msg=2000GeV.png}
\caption{Dependence of the $K$-factor for the process $uu \to \tilde{u}\tilde{u}$ on the mass of the pseudoscalar sgluon. For $m_{\sigma} \approx \unit[10^5]{GeV}$ onwards want find a logarithmic scaling as predicted by \cite{Cheng:1997sq}.}\label{fig:SgluonMassDependence}
\end{center}
\end{figure}
This is actually expected as an effective field theory of the MRSSM where the sgluon is integrated out is no longer supersymmetric. This is because the sgluon is together with the octino part of a supermultiplet. Integrating out only the sgluon means that the octino misses its superpartner in the effective field theory. In this case the decoupling theorem \cite{Appelquist:1974tg} does no longer hold\footnote{It is even possible to predict the  logarithmic scaling of $\sigma(qq \to \tilde{q}\tilde{q})$ quantitatively. The coefficient of the logarithm is proportional to the difference of the one-loop $\beta$ function coefficients of $g_s$ and $\hat{g}_s$ in the effective theory with the heavy particle integrated out, see eq. (4) in \cite{Cheng:1997sq}.}.
The logarithmic scaling of the cross section as a function of the mass of the pseudoscalar is shown in fig. \ref{fig:SgluonMassDependence}.

>>>>

\subsection{The Squark Production Cross Section at Next-to-Leading Order}
The squark production cross section at next-to-leading order is composed out of three contributions. These are the Born (tree-level) cross section, the virtual and the real corrections. 
\begin{align}
\sigma^{\mathrm{NLO}} = \sigma^{\mathrm{B}} + \sigma^{\mathrm{V}} + \sigma^{\mathrm{R}}\label{eq:BVR}
\end{align}
All three contributions have to be calculated with next-to-leading order parton density functions and $\alpha_s$. There are seven mass scales: the squark mass $m_{\tilde{q}}$, the gluino mass $m_{\tilde{g}}$, the center of mass energy of the colliding protons which is set to $\sqrt{S} = \unit[13]{TeV}$, the top-quark mass which is fixed to $m_t = \unit[172]{GeV}$, the mass of the pseudoscalar $m_{\sigma}$ and the renormalization and factorization scale. They have been set to $\mu_F = \mu_R = m_{\tilde{q}}$. The next section summaries how the calculation has been automated within a \texttt{C++} program.

@@

\subsubsection{Remark on Prefactors in the Calculation}\label{sec:prefactors}

@@

\caption{Dependence of the $K$-factor for the process $pp \to \tilde{u}_L\tilde{u}_R$ in the MRSSM on the squark mass $m_{\tilde{q}}$ for fixed $m_{\tilde{g}} = \unit[2000]{GeV}$ (top) and on the gluino mass $m_{\tilde{g}}$ for fixed $m_{\tilde{q}} = \unit[1000]{GeV}$ (bottom). The mass of the pseudoscalar is $m_{\sigma} = \unit[5000]{GeV}$. The used parton density function are the one from table \ref{tab:MRSSMKfactors}. The missing smoothness of the plot originates from the uncertainty of the integration.}\label{fig:1LXsection_fixed_m_MRSSM}

@@

\caption{$K$-factors of the process $pp \to \tilde{u}_L\tilde{u}_R$  in the MRSSM for a selected set of masses. The pseudoscalar mass is fixed to \mbox{$m_{\mathrm{\sigma}} = \unit[5000]{GeV}$}. For $\sigma^{\mathrm{LO}}$ parton density functions \texttt{MMHT2014LO} (LHAPDF ID: $25000$) is used, whereas for $\sigma^{\mathrm{NLO}}$ \texttt{MMHT2014NLO} (LHAPDF ID: $25100$) is used.}\label{tab:MRSSMKfactors}

@@

It can be seen that the $K$-factors depend strongly on the ratio $\frac{m_{\tilde{q}}}{m_{\tilde{g}}}$ but not that strongly on the masses itself. This is also depicted in fig. \ref{fig:Kfactor_MRSSM} which shows the $K$-factors in the $m_{\tilde{q}}$-$m_{\tilde{g}}$ plane. The contour lines follow approximately rays from the origin. By performing the calculation without the shuffling of prefactors as explained in section \ref{sec:prefactors}, the raising of the $K$-factors for increasing values of $\frac{m_{\tilde{q}}}{m_{\tilde{g}}}$ has been traced down to the real corrections. As quark radiation is not included, this effect can further be pinned down to originate from the diagrams in fig. \ref{fig:RealGluon}. However, at least for small $m_{\tilde{g}}$, fig. \ref{fig:Kfactor_MRSSM} suggests that the $K$-factors drop again for increasing $m_{\tilde{q}}$.
\includegraphics[scale=.6]{figures/Kfactor_MRSSM.png}
\caption{Contour plot of the $K$-factor of up-squark production in the MRSSM. The pseudoscalar mass is set to $m_{\sigma} = \unit[5000]{GeV}$. The used parton density functions are the same as in table \ref{tab:MRSSMKfactors}.}\label{fig:Kfactor_MRSSM}

@@

\subsubsection{$K$-factors for Squark Production in the MSSM}
To compare the results of $K$-factors in the MRSSM to those in the MSSM, the cross sections in question have been calculated using the program \texttt{MadGraph5\_aMC@NLO} with the same parton density functions as before, i.e. \texttt{MMHT2014}. When generating the code for the pertaining processes, e.g. two protons to $\tilde{u}_L + \tilde{u}_R$, with
\begin{lstlisting}[style=Mybash]
generate p p > ul ur [QCD] $$go
\end{lstlisting}
the $s$-channel gluinos were omitted, with the addendum \texttt{\$\$go} as \texttt{MadGraph5} does not deal with divergences coming from the gluino in the on-shell limit. 
\begin{figure}[!htpb]
\begin{center}
\includegraphics[scale=.46]{figures/MSSM_uu_susu_Kfactors_msg=2000GeV.png}
\includegraphics[scale=.46]{figures/MSSM_uu_susu_Kfactors_msq=1000GeV.png}
\caption{Dependence of the $K$-factor for the process $pp \to \tilde{u}\tilde{u}$, where $\tilde{u}\tilde{u} \in \left\{ \tilde{u}_L\tilde{u}_R, \tilde{u}_L\tilde{u}_L, \tilde{u}_R\tilde{u}_R \right\}$, in the MSSM on the squark mass $m_{\tilde{q}}$ for fixed $m_{\tilde{g}} = \unit[2000]{GeV}$ (top) and on the gluino mass $m_{\tilde{g}}$ for fixed $m_{\tilde{q}} = \unit[1000]{GeV}$ (bottom). The used parton density functions are the same as in table \ref{tab:MRSSMKfactors}. The $K$-factors have been broken down to contributions coming from $pp \to \tilde{u}_L\tilde{u}_R$ and from $pp \to \tilde{u}_L\tilde{u}_L$. The latter equals the contributions coming from $pp \to \tilde{u}_R\tilde{u}_R$. The missing smoothness of the plot originates from the uncertainty of the integration.}\label{fig:1LXsection_fixed_m_MSSM}
\end{center}
\end{figure}
As already discussed, these contributions are only of sub percent order when $m_{\tilde{q}} > m_{\tilde{g}}$\cite{Gavin:2013kga} but increase for $m_{\tilde{g}} > m_{\tilde{q}}$ as indicated in table \ref{tab:MSSMMadGraphProspino}, which compares $\sigma^{\mathrm{NLO}}$ obtained using \texttt{MadGraph5} and the program \texttt{Prospino2}\cite{Beenakker:1996ed} for both scenarios.

\begin{table}[H]
\begin{center}
\begin{tabular}{c|c?c?c}
$m_{\tilde{q}}$ in GeV & $m_{\tilde{g}}$ in GeV & $\sigma^{\mathrm{NLO}}_{\texttt{MadGraph}}$ in fb & $\sigma^{\mathrm{NLO}}_{\texttt{Prospino2}}$ in fb \\
\hlinewd{2pt}
$1500$ & $1000$ & $9.49 \pm 0.03$ & $9.48 \pm 0.09$ \\
$1000$ & $2000$ & $13.32 \pm 0.05$ & $13.89 \pm 0.11$ 
\end{tabular}
\caption{$\sigma^{\mathrm{NLO}}(pp \to \tilde{u}_L\tilde{u}_R)$ in the MSSM obtained from \texttt{MadGraph5} and \texttt{Prospino2} for a selected set of masses. \texttt{MadGraph5} uses parton density function given in \ref{tab:MRSSMKfactors} whereas \texttt{Prospino2} uses \texttt{CTEQ6L1} parton density functions\cite{Pumplin:2002vw} for leading order calculations and \texttt{CTEQ66}\cite{Nadolsky:2008zw} for next-to-leading order calculations. The stated uncertainties originate from the phase space integration. Within the \texttt{MadGraph5} calculation, the $s$-channel gluino diagrams have been discarded. This becomes relevant only when $m_{\tilde{g}} > m_{\tilde{q}}$.}\label{tab:MSSMMadGraphProspino}
\end{center}
\end{table}

@@

\begin{table}[H]
\begin{center}
\begin{tabular}{c|c?c|c|c?c|c|c}
\multicolumn{2}{c?}{} & \multicolumn{3}{c?}{$p + p \to \tilde{u}_L + \tilde{u}_R$} & \multicolumn{3}{c}{$p + p \to \tilde{u} + \tilde{u}$} \\
\hlinewd{2pt}
$m_{\tilde{q}}$ in GeV & $m_{\tilde{g}}$ in GeV & $\sigma^{\mathrm{LO}}$ in fb & $\sigma^{\mathrm{NLO}}$ in fb & $K$ & $\sigma^{\mathrm{LO}}$ in fb & $\sigma^{\mathrm{NLO}}$ in fb & $K$\\
\hlinewd{2pt}
$500$ & $500$ & $966.4$ & $1328$ & $1.37$ & $2333$ & $3102$ \ignore{1328+2*887} & $1.33$\\
$500$ & $1000$ & $303.4$ & $352$ & $1.16$ & $1271$ & $1393$ \ignore{353 + 2*520} & $1.10$\\
$1000$ & $1000$ & $42.63$ & $58.5$ & $1.37$ & $122.2$ & $161.3$ \ignore{58.5+2*51.4} & $1.32$\\
$1000$ & $2000$ & $11.56$ & $13.3$ & $1.15$ & $66.36$ & $72.1$ \ignore{13.3+2*29.4} & $1.09$\\
$1500$ & $3000$ & $0.9166$ & $1.05$ & $1.14$ & $7.045$ & $7.57$\ignore{1.05+2*3.26} & $1.07$
\end{tabular}
\caption{$K$-factors of the process $pp \to \tilde{u}_L\tilde{u}_R$ and $pp \to \tilde{u}\tilde{u}$, where \mbox{$\tilde{u}\tilde{u} \in \left\{ \tilde{u}_L\tilde{u}_R, \tilde{u}_L\tilde{u}_L, \tilde{u}_R\tilde{u}_R \right\}$,} in the MSSM for a selected set of masses. The used parton density functions are the same as in table \ref{tab:MRSSMKfactors}.}\label{tab:KfactorsMSSM}
\end{center}
\end{table}
\begin{figure}[H]
\begin{center}
\includegraphics[scale=.6]{figures/Kfactor_MSSM.png}
\caption{Contour plot of the $K$-factor of the process  $pp \to \tilde{u}_L\tilde{u}_R$ and $pp \to \tilde{u}\tilde{u}$, where \mbox{$\tilde{u}\tilde{u} \in \left\{ \tilde{u}_L\tilde{u}_R, \tilde{u}_L\tilde{u}_L, \tilde{u}_R\tilde{u}_R \right\}$,} in the MSSM. The used parton density functions are the same as in table \ref{tab:MRSSMKfactors}.}\label{fig:Kfactor_MSSM}
\end{center}
\end{figure}
Again, it is clearly visible that the $K$-factors depend rather strongly on the ratio $\frac{m_{\mathrm{g}}}{m_{\mathrm{q}}}$ but not so much on the masses itself (see also fig. \ref{fig:Kfactor_MSSM}). As the real corrections are identical to those in the MRSSM, it is again these very real corrections which are responsible for an enhancement of the $K$-factor for large values of  $\frac{m_{\mathrm{g}}}{m_{\mathrm{q}}}$.
Comparing with the results from the MRSSM in table \ref{tab:MRSSMKfactors}, one can state that the $K$-factors of the MRSSM are slightly larger. This might be the case, as there are additional degrees of freedom in the $R$-symmetric model, namely the octino which enters the gluino self-energy contributions (see second diagram from the right in \ref{fig:GluinoSE}) and sgluons which enter box diagrams (see lower right diagram in \ref{fig:1loopdiagrams}). The dependence on the sgluon mass is discussed within the next section.\\
The uncertainties of next-to-leading order cross sections, due to the phase space integration, within this section are of percent order. The uncertainty which corresponds to the scale variation does not increase 15\%.

\subsection{The Cross Section in the Limit of Large Sgluon Masses}
\begin{figure}[H]
\begin{center}
\includegraphics[scale=.5]{figures/MRSSM_uu_susu_Kfactors_msq=1000GeV_msg=2000GeV.png}
\caption{Dependence of the $K$-factor for the process $pp \to \tilde{u}\tilde{u}$ on the mass of the pseudoscalar sgluon. Squark and gluino mass have been fixed to $m_{\tilde{q}} = \unit[1000]{GeV}$ and $m_{\tilde{g}} = \unit[2000]{GeV}$ . The used parton density functions are the same as in table \ref{tab:MRSSMKfactors}. For $m_{\sigma} \approx \unit[10^5]{GeV}$ onwards, one finds a logarithmic scaling as predicted by \cite{Cheng:1997sq}.}\label{fig:SgluonMassDependence}
\end{center}
\end{figure}
The cross section for squark production does not exist in the limit of an infinitely large sgluon mass, instead it was found that it diverges logarithmically:
\begin{align}
\lim_{m_{\sigma}\to\infty} \sigma(qq \to \tilde{q}\tilde{q}) \sim \ln \frac{m_{\sigma}^2}{\mu^2}.
\end{align}
This is actually expected as an effective field theory of the MRSSM where the sgluon is integrated out is no longer supersymmetric. This is because the sgluon is together with the octino part of a supermultiplet. Integrating out only the sgluon means that the octino misses its superpartner in the effective field theory. In this case the decoupling theorem \cite{Appelquist:1974tg} does no longer hold\footnote{It is even possible to predict the  logarithmic scaling of $\sigma(qq \to \tilde{q}\tilde{q})$ quantitatively. The coefficient of the logarithm is proportional to the difference of the one-loop $\beta$ function coefficients of $g_s$ and $\hat{g}_s$ in the effective theory with the heavy particle integrated out, see eq. (4) in \cite{Cheng:1997sq}.}.
The logarithmic scaling of the cross section as a function of the mass of the pseudoscalar is shown in fig. \ref{fig:SgluonMassDependence}.
